---
title: "Colugo flight path analysis"
author: "Rachel Lee S.K. "
date created: "06/03/2021"
date updated: "13/11/2022"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

# Setup packages and working directories

```{r Load libraries and setup WD, include = F}
library(tidyverse)
library(raster)
library(rgdal)
library(units)
library(sf)
library(ggplot2)
library(tmap)
library(terra)

setwd("~/Documents/Manuscripts/Colugo MS")

```

# Flight path generation (2016)

Flight path generation between each pairs of tree (before hoarding i.e. 2016)

```{r Join attributes by location & line generation, include = F}

#variable control panel
path_trees2016 <- "./Working layers/Trees_50mMLR_2016_P5P6MLR.shp" #Trees_2016_updated.shp

#load & inspect data
trees16_sf <- st_read(path_trees2016)
head(trees16_sf)
unique(trees16_sf$MLRLoc)
unique(trees16_sf$Species)

#exclude plants that are unsuitable for gliding mammal usage e.g., shrubs or spiny palms
genus_exclude <- 'Elaeis|Oncosperma|Strychnos|Ziziphus|Clerodendrum|Dillenia|Calliandra' 


#data transformation
trees16_sf2<- trees16_sf %>%
  mutate(Geom = trees16_sf$geometry) %>%
  mutate_at("MLRLoc", as.character) %>% 
  filter(! grepl(genus_exclude, Species)) %>% ##remove unsuitable genus
  filter(! Height_M < 1) #remove short individuals

#generate unique buffer based on estimated glide distance
trees16_buffer <- st_buffer(trees16_sf2, trees16_sf2$GlideRadii)

#join attributes by location
#join ToTrees and FrmTrees within glide radii
#remove those from same sides (e.g. North to North)
#remove those from South to S - Behind NS Fence South
joined_trees16 <- st_join(trees16_sf2, trees16_buffer, 
                          suffix = c(".To", ".Frm"), left = FALSE) %>% ##To before frm
                  filter(! MLRLoc.Frm == MLRLoc.To) %>% 
                  filter(! grepl("S", MLRLoc.Frm) == grepl("S", MLRLoc.To))

#line generation
flightlines <- st_sfc(mapply(function(a,b){st_cast(st_union(a,b),"LINESTRING")},
                                        joined_trees16$Geom.Frm, joined_trees16$Geom.To,
                                        SIMPLIFY=FALSE),
                      crs=3414)
head(flightlines) #does not contain attributes

#random check to ensure that order remained the same
flightlines[[211]]
joined_trees16[211, c(24,25)]

#Updating geometries of joined_trees16 from multipoint to multiline
#allows the retention of attributes
flight16 <- joined_trees16 %>%
  st_set_geometry(flightlines) %>%
  dplyr::select(-c(Lat.Frm, Long.Frm, Height_GR.Frm, Geom.Frm,
                   Lat.To, Long.To, Height_GR.To, Geom.To))

#check crs
st_crs(flightlines)

```

####Set landing height (2016)

By default, landing height of Colugo is set to 1.5m above ground. However, if there is a hump (rise in elevation) along the flight path, the Colugo might risk hitting the hump if it were "forced" to land at 1.5m for the landing tree. Hence, the landing height has to be adjusted in such scenarios.

1. Determine the max elevation value along the flight path from the raster digital elevation model
2. Find difference between maxElev and Elev.To; colname: `diffElev_Max`
3. If diffElev_Max > 1.5, set the landing height (`LandHeight`) as the difference. If not, landing height remains as 1.5m

```{r Extract data from raster, include = F}

#variable control panel
path_dem2016 <- "./Working layers/DEM_2016_crop.tif"
path_MLRbuffer <- "./Working layers/P5P6_FutureMLR_50mBuffer.shp"

#load data
dem16_Crop <- rast(path_dem2016)
  
mlr_buffer <- st_read(path_MLRbuffer)

tm_shape(dem16_Crop) +
  tm_raster() +
tm_shape(mlr_buffer) +
  tm_polygons(col = "grey", alpha = 0.5)

#### Measure raster values ####
#extract max raster value along each flight path
flightSamp <- flight16 %>%
  mutate(maxElev = terra::extract(x = dem16_Crop, y = vect(flight16), fun = max)$layer)
  # mutate(maxElev = raster::extract(x = dem16_Crop, y = ., fun = max)) #old code using raster package

#### Set landing height for F1 ####
flight16_dem2  <- flightSamp %>%
  mutate(diffElev_Max = maxElev-Elev.To) %>%
  mutate(LandHeight = ifelse(diffElev_Max>1.5, diffElev_Max, 1.5))

#### Save out shapefile ####
st_write(flight16_dem2, "./Results/Flight paths/FlightPaths_2016_WElev.shp", crs=3414)

```

####Clear tree height, no hoard

1. Check that land tree is tall enough for Colugo to land: `0.8*Height_M.Frm > LandHeight`
2. Measure minimum take-off height (H) which is what's required for the Colugo to travel along each flight path and reach respectiveLandHeight: `H = v + LandHeight + (Elev.To - Elev.Frm)`
3. Check if the departure tree is tall enough for the Colugo to land at the paired landing tree: `0.8*Height_M.Frm >  H`
4. Find highest point where Colugo can depart for each flight path: `MaxDepart = v + 0.8*Height_M.To + (Elev.To - Elev.Frm)`


```{r Removing invalid flight paths, include = F}

#variable control panel
path_flight16 <- "./Results/Flight paths/FlightPaths_2016_WElev.shp"

#load & inspect data
flight16 <- st_read(path_flight16)
colnames(flight16)

#st_write process changed the col names
# shapefile attributes rename
flight16 <- flight16 %>%
  rename(Height_M.Frm = Hgh_M_F) %>%
  rename(Height_M.To = Hgh_M_T) %>%
  rename(LandHeight = LndHght) %>%
  rename(Elev.To = Elev_To) %>%
  rename(Elev.Frm = Elv_Frm)  

#### Calculate H ####
flight16 <- flight16 %>%
  #measure horizontal glide distance
  mutate(Dist = drop_units(st_length(flight16))) %>%
  #determine change in vertical height through colugo glide ratio model
  mutate(v = Dist/(0.8781*log(Dist) + 0.046)) %>%
  #determine minimum departure height with consideration of elevation
  mutate(H = v + LandHeight + (Elev.To - Elev.Frm)) %>%
  #set upper bound for departure height
  mutate(MaxDepart = v + 0.8*Height_M.To + (Elev.To - Elev.Frm)) %>% 
  mutate(Depart = ifelse((MaxDepart<0.8*Height_M.Frm), MaxDepart, 0.8*Height_M.Frm)) %>%
  #find DiffHij, which is used to check path feasibility and used as path quality
  mutate(Diff_H = Depart-H ) %>%
  #check also that landing tree is tall enough to support colugo landing
  mutate(Diff_M = 0.8*Height_M.To - LandHeight)

#check again that unsuitable genus is not present
flight16_plantrem <- flight16 %>%
  filter(! grepl(genus_exclude, Specs_T)) %>%
  filter(! grepl(genus_exclude, Spcs_Fr))

st_write(flight16_plantrem, "./Results/Flight paths/FlightPaths_2016_with Height.shp", crs=3414)

#### Remove invalid flight paths ####
flight16_valid <- flight16_plantrem %>% 
  filter(! Diff_H < 0) %>%
  filter(! Diff_M < 0)

st_write(flight16_valid, "./Results/Flight paths/FlightPaths_2016_Valid.shp", crs=3414)

#examines differences in connectivity with and without Alstonia (planted roadside trees)
flight16_noAlstonia <- flight16_valid %>%
  filter(! grepl("Alstonia", Specs_T)) %>%
  filter(! grepl("Alstonia", Spcs_Fr))
flight16_onlyAlstonia <- flight16_valid %>%
  filter(grepl("Alstonia", Specs_T) | grepl("Alstonia", Spcs_Fr))

#plot to check  
tm_shape(flight16_valid) +
    tm_lines(col = 'gray', alpha = 0.4, lwd = 1.5, legend.show = FALSE) +
tm_shape(flight16_onlyAlstonia) +
    tm_lines(col = 'steelblue', alpha = 0.4, lwd = 1.5, legend.show = FALSE)

```

####Rasterize 2016 flight paths

For visualization and to calculate percentage pixels covered by feasible glide paths

```{r Rasterize, include = F}

#if load from saved data
path_flight16 <- "./Results/Flight paths/FlightPaths_2016_Valid.shp"
path_MLRbuffer <- "./Working layers/P5P6_FutureMLR_50mBuffer.shp"
flight16_valid <- st_read(path_flight16)
mlr_buffer <- st_read(path_MLRbuffer)

#if load from previous code
#i.e., "flight16_valid"

head(flight16_valid)

#Dilation to allow for variance in the linearity of a colugo glide path 
#as colugos have some manoeuvrability in mid-air and can glide around minor obstacles
flight16_valid_buffer <- st_buffer(flight16_valid, 2, endCapStyle = 'FLAT')

#separate into south to north
#and north to south
s2n_16 <- flight16_valid_buffer %>%
  filter(MLRLc_T == "North")
n2s_16 <- flight16_valid_buffer %>%
  filter(MLRLc_T == "South")

#rasterize each direction first
raster_template = raster(extent(mlr_buffer), resolution = 0.5,
                         crs = st_crs(mlr_buffer)$proj4string)

flight16_raster_s2n = rasterize(s2n_16, raster_template, field = 'Diff_H', fun = 'count', progress = 'text')
flight16_raster_n2s = rasterize(n2s_16, raster_template, field = 'Diff_H', fun = 'count', progress = 'text')

#rasterize both direction
flight16_raster_total = rasterize(flight16_valid_buffer, raster_template, field = 'Diff_H', fun = 'count', progress = 'text')

#rasterize based on quality of flight
flight16_raster_quality = rasterize(flight16_valid_buffer, raster_template, 
                                    field = 'Diff_H', fun = 'sum', 
                                    progress = 'text')

#find min of values for overlapping cells (effective connectivity)
rasterlist_16 <- list(flight16_raster_n2s,flight16_raster_s2n )
stack16 <- stack(rasterlist_16)

minConnectivity_16 <- min(stack16)

#plot and see
plot(flight16_raster_quality)
plot(flight16_raster_total)
plot(flight16_raster_s2n)
plot(flight16_raster_n2s)

plot(minConnectivity_16)

#save out
writeRaster(flight16_raster_total, "./Results/Flight paths/FlightPaths_2016_Valid_raster.tif", crs=3414)
writeRaster(flight16_raster_s2n, "./Results/Flight paths/FlightPaths_2016_Valid_raster_s2n.tif", crs=3414)
writeRaster(flight16_raster_n2s, "./Results/Flight paths/FlightPaths_2016_Valid_raster_n2s.tif", crs=3414)
# writeRaster(minConnectivity_16, "./Results/Flight paths/FlightPaths_2016_Valid_raster_effective connectivity.tif", crs=3414)

```

####2019
####Preparation for hoarding

<skip to chunk 6 after flight path generated>

Files to be prepared in QGIS: 
1. Clipped hoarding line shapefile that contains South/North information
2. Hoarding polygon shapefile that contains South/North information

Files to be prepared in R:
1. Flight path from 2019 trees and poles

```{r Join attributes by location & line generation, include = F}

#variable control panel
path_trees2019 <- "./Working layers/TreesPoles_Remain2019_P5P6MLR_20210604.shp" #TreesPoles_2019_updated.shp
genus_exclude <- 'Elaeis|Oncosperma|Strychnos|Ziziphus|Clerodendrum|Dillenia|Calliandra'

#load tree data
trees19_sf <- st_read(path_trees2019)

#data transformation
trees19_sf2<- trees19_sf %>%
  mutate(Geom = trees19_sf$geometry) %>%
  mutate_at("MLR_Loc", as.character) %>%
  filter(! grepl(genus_exclude, Species)) %>%
  filter(! Height_M < 1)

trees19_buffer <- st_buffer(trees19_sf2, trees19_sf2$GlideRadii)

#join attributes by location
#separate into poles and trees
#for poles, allow for same side flight i.e. S to S and N to N
#required to check multistep travel from S to Pole to N and vice versa
joined_trees19 <- st_join(trees19_sf2, trees19_buffer, 
                          suffix = c(".To", ".Frm"), left = FALSE) %>%
                  filter(! MLR_Loc.Frm == MLR_Loc.To) %>% 
                  filter(! grepl("S", MLR_Loc.Frm) == grepl("S", MLR_Loc.To))

################for future scenario (no hoarding, no poles)#####################
# #recylcing of codes to be used
# #turn on the following chunk when generating this scenario.
# #the main difference is that...
# #poles removed from tree layer before joining attributes by location
# trees22 <- trees19_sf2 %>%
#  filter(Type == 'Tree')
# 
# trees22_buffer <- st_buffer(trees22, trees22$GlideRadii)
# 
# joined_trees19 <- st_join(trees22, trees22_buffer,
#                          suffix = c(".To", ".Frm"), left = FALSE) %>%
#                  filter(! MLR_Loc.Frm == MLR_Loc.To)

################for future scenario (no hoarding, no poles)#####################

#line generation
flightlines19 <- st_sfc(mapply(function(a,b){st_cast(st_union(a,b),"LINESTRING")},
                                        joined_trees19$Geom.Frm, joined_trees19$Geom.To,
                                        SIMPLIFY=FALSE),
                      crs=3414)
head(flightlines19) #does not contain attributes

#random check to ensure that order remained the same
flightlines19[[211]]
joined_trees19[211, c(24,25)]

#### Flight path generation ####
flight19 <- joined_trees19 %>%
  st_set_geometry(flightlines19) %>%
  dplyr::select(-c(Lat.Frm, Long.Frm, Height_GR.Frm, Geom.Frm,
                   Lat.To, Long.To, Height_GR.To, Geom.To))

flightcolumns <- colnames(flight19)

#### Save shapefile ####
st_write(flight19, "./Results/Flight paths/FlightPaths_2019.shp")
#write.csv(flightcolumns, "./R/Flight_colnames_without Elev.csv")

#For future with no hoarding and no poles
# st_write(flight19, "./Results/Flight paths/FlightPaths_2022.shp")

```

####Set landing height (2019)

By default, landing height of Colugo is set to 1.5m above ground. However, if there is a hump (rise in elevation) along the flight path, the Colugo might risk hitting the hump if it were "forced" to land at 1.5m for the landing tree. Hence, the landing height has to be adjusted in such scenarios.

1. Determine the elevation value along the flight path from the raster digital elevation model
2. Find difference between maxElev and Elev.To; colname: `diffElev_Max`
3. If diffElev_Max > 1.5, set the landing height (`LandHeight`) as the difference. If not, landing height remains as 1.5m


```{r Extract data from raster, include = F}

#variable control panel
path_dem2019 <- "./Working layers/DEM_2019_crop.tif"
path_MLRbuffer <- "./Working layers/P5P6_FutureMLR_50mBuffer.shp"
path_flight19 <- "./Results/Flight paths/FlightPaths_2019.shp"
# path_flight22 <- "./Results/Flight paths/FlightPaths_2022.shp"

#load data
dem19_Crop <- rast(path_dem2019)
# dem19_Crop <- raster(path_dem2019) #old code that doesn't run anymore
mlr_buffer <- st_read(path_MLRbuffer)
flight19 <- st_read(path_flight19)
# flight19 <- st_read(path_flight22)

tm_shape(dem19_Crop) +
  tm_raster() +
tm_shape(mlr_buffer) +
  tm_polygons(col = "grey", alpha = 0.5)

#### Measure raster values ####
flightSamp <- flight19  %>%
  mutate(maxElev = terra::extract(x = dem19_Crop, y = vect(flight19), fun = max)$DEM_2019)

# flightSamp <- flight19 %>% #old code that doesn't run anymore
#   mutate(maxElev = raster::extract(x = dem19_Crop, y = flight19, fun = max)$DEM_2019)

#### Set landing height for F1 ####
flightSamp <- flightSamp %>%
  mutate(diffElev_Max = maxElev-Elev_To) %>%
  mutate(LandHeight = ifelse(diffElev_Max>1.5, diffElev_Max, 1.5))

#### Save out shapefile ####
#skip save if running for 2022
st_write(flightSamp, "./Results/Flight paths/FlightPaths_2019_WElev.shp", crs=3414)

```

#### Clear tree height WITHOUT hoarding and poles for 2022

```{r Removing invalid flight paths for simulation, include = F}

#### Calculate H ####
flight22 <- flightSamp %>%
  mutate(Dist = drop_units(st_length(flight19))) %>%
  mutate(v = Dist/(0.8781*log(Dist) + 0.046)) %>%
  mutate(H = v + LandHeight + (Elev_To - Elv_Frm)) %>%
  mutate(MaxDepart = v + 0.8*Hgh_M_T + (Elev_To - Elv_Frm)) %>% 
  mutate(Depart = ifelse((MaxDepart<0.8*Hgh_M_F), MaxDepart, 0.8*Hgh_M_F)) %>%
  mutate(Diff_H = Depart-H ) %>%
  mutate(Diff_M = 0.8*Hgh_M_T - LandHeight)

#### Remove invalid flight paths ####
flight22_valid <- flight22 %>% 
  filter(! Diff_H < 0) %>%
  filter(! Diff_M < 0)

# st_write(flight22_valid, "./Results/Flight paths/FlightPaths_2022_Valid.shp", crs=3414)

#plot to check
tm_shape(flight22_valid) +
  tm_lines(col = 'blue', alpha = 1, lwd = 1.5, legend.show = FALSE)

```


#### Clipping flight paths with hoarding (2019)

The consideration of hoarding as obstacles has to be unidirectional i.e. North to South and South to North to be considered separetly.

Need to generate 3 sets of flight path where:
1. *F1*: ends at landing tree
2. *F2*: ends at 0.5m above hoarding at landing side
3. *F3*: ends at hoarding at departure side
2. *F4*: ends at 1.0m above hoarding at landing side

```{r Join attributes by location & difference, include = F}

#variable control panel
path_flight19 <- "./Results/Flight paths/FlightPaths_2019_WElev.shp"
path_colnames <- "./R/Flight_colnames_without Elev.csv"

path_hoardLines <- "./Working layers/P5P6_Hoard_WElev_Jan2019.shp"
path_hoardPoly <- "./Working layers/Hoard_WElev_Jan2019_Poly.shp"

#load data
hoard <- st_read(path_hoardLines)
hoardPoly <- st_read(path_hoardPoly)

#if continuing from previous code chunk
#flight19 <- flightSamp

#if starting with new r environment
flight19 <- st_read(path_flight19)
flightcolumns <- read.csv(path_colnames)

#### update flight19 column names that were changed during st_write ####
colnames(flight19)

flightcolumns2 <- flightcolumns[1:16,] %>%
  dplyr::select(x) %>%
  add_row(x=c("maxElev", "diffElev_Max", "LandHeight", "geometry"))

flight19 <- rename_with(flight19, function(x) flightcolumns2$x)

#### start ####
#separate North South
F1_n2s <- flight19 %>%
  filter(MLR_Loc.Frm == 'North')

F1_s2n <- flight19 %>%
  filter(! MLR_Loc.Frm == 'North')

# Continue here for pole continuty check #
# variable control panel (for sides)
#run one at a time (n2s or s2n)
#### north to south ####
flight19_run <- F1_n2s
F2side_exclude <- 'North'
F3side_exclude <- 'South'

#add elevation data where hoard intersect flight
joined_F2 <- st_join(flight19_run, hoard,
                             left = FALSE) %>%
                     mutate_at("Side", as.character) %>%
                     filter(! grepl(F2side_exclude, Side))
joined_F3 <- st_join(flight19_run, hoard,
                             left = FALSE) %>%
                     mutate_at("Side", as.character) %>%
                     filter(! grepl(F3side_exclude, Side))

#clip away section beyond hoarding
F2_n2s <- st_difference(joined_F2, 
                        hoardPoly[which(hoardPoly$Side == F3side_exclude),]) %>%
  mutate(LandHeight1 = 2.9) %>%
  mutate(LandHeight2 = 3.4)
F3_n2s <- st_intersection(joined_F3, 
                  hoardPoly[which(hoardPoly$Side == F2side_exclude),]) %>%
  mutate(LandHeight = 2.9)

#plot to check
tm_shape(hoard) +
    tm_lines(col = 'black') +
tm_shape(F1_n2s) +
    tm_lines(col = 'gray', lwd = 1.5, legend.show = FALSE) +
tm_shape(F2_n2s) +
    tm_lines(col = 'steelblue', lwd = 1.5, legend.show = FALSE) +
tm_shape(F3_n2s) +
    tm_lines(col = 'indianred')

#variable control panel (for sides)
#run one at a time (n2s or s2n)
#### south to north ####
flight19_run <- F1_s2n
F2side_exclude <- 'South'
F3side_exclude <- 'North'

#add elevation data where hoard intersect flight
joined_F2 <- st_join(flight19_run, hoard,
                             left = FALSE) %>%
                     mutate_at("Side", as.character) %>%
                     filter(! grepl(F2side_exclude, Side))
joined_F3 <- st_join(flight19_run, hoard,
                             left = FALSE) %>%
                     mutate_at("Side", as.character) %>%
                     filter(! grepl(F3side_exclude, Side))

#clip away section beyond hoarding
F2_s2n <- st_difference(joined_F2, 
                        hoardPoly[which(hoardPoly$Side == F3side_exclude),]) %>%
  mutate(LandHeight1 = 2.9) %>%
  mutate(LandHeight2 = 3.4)
F3_s2n <- st_intersection(joined_F3, 
                  hoardPoly[which(hoardPoly$Side == F2side_exclude),]) %>%
  mutate(LandHeight = 2.9)

#plot to check
tm_shape(hoard) +
    tm_lines(col = 'black') +
tm_shape(F1_s2n) +
    tm_lines(col = 'gray', lwd = 1.5, legend.show = FALSE) +
tm_shape(F2_s2n) +
    tm_lines(col = 'steelblue', lwd = 1.5, legend.show = FALSE) +
tm_shape(F3_s2n) +
    tm_lines(col = 'indianred')

```

####Clear tree height, with hoarding

1. Check that land tree is tall enough for Colugo to land: `0.8*Height_M.Frm > LandHeight`
2. Measure minimum take-off height (H) which is what's required for the Colugo to travel along each flight path and reach respectiveLandHeight: `H = v + LandHeight + (Elev.To - Elev.Frm)`
3. Find maximum H value *maxH*
4. Check if the departure tree is tall enough for the Colugo to land at the paired landing tree: `0.8*Height_M.Frm >  maxH`
4. Find highest point where Colugo can depart for each flight path: `MaxDepart = v + 0.8*Height_M.To + (Elev.To - Elev.Frm)`


```{r Removing invalid flight paths, include = F}

#### variable control panel (for sides) #### 
#run one at a time (n2s or s2n)
#north to south
F1_run <- F1_n2s
F2_run <- F2_n2s
F3_run <- F3_n2s

#south to north
F1_run <- F1_s2n
F2_run <- F2_s2n
F3_run <- F3_s2n

#### Calculate H ####
#run one at a time (n2s or s2n)
#to calculate for each flight resulting in
#H1, H2, H3, H4
#find maxH2 between H1, H2 and H3
#find maxH4 between H1, H4 and H3
  
F2_df <- F2_run %>%
  mutate(Dist = drop_units(st_length(F2_run))) %>%
  mutate(v2 = Dist/(0.8781*log(Dist) + 0.046)) %>%
  mutate(H2 = v2 + LandHeight1 + (Elevation - Elev.Frm)) %>% #rmb to set To Elevation to the elevation of hoarding
  mutate(H4 = v2 + LandHeight2 + (Elevation - Elev.Frm)) %>% #rmb to set To Elevation to the elevation of hoarding
  st_drop_geometry() %>%
  dplyr::select(c("ID.To", "ID.Frm", "H2", "H4"))

#duplicated <- F2_df[duplicated(F2_df[,c(1,9)]),]
  
F3_df <- F3_run %>%
  mutate(Dist = drop_units(st_length(F3_run))) %>%
  mutate(v3 = Dist/(0.8781*log(Dist) + 0.046)) %>%
  mutate(H3 = v3 + LandHeight + (Elevation - Elev.Frm)) %>% #rmb to set To Elevation to the elevation of hoarding
  st_drop_geometry() %>%
  dplyr::select(c("ID.To", "ID.Frm", "H3"))
  
master <- F1_run %>%
  mutate(Dist = drop_units(st_length(F1_run))) %>%
  mutate(v = Dist/(0.8781*log(Dist) + 0.046)) %>%
  mutate(H1 = v + LandHeight + (Elev.To - Elev.Frm)) %>%
  left_join(F2_df, by = c("ID.To", "ID.Frm"), keep = F) %>%
  left_join(F3_df, by = c("ID.To", "ID.Frm"), keep = F) %>%
  rowwise() %>%
  mutate(maxH2 = max(H1, H2, H3, na.rm = T)) %>%
  mutate(maxH4 = max(H1, H4, H3, na.rm = T)) %>%
  mutate(M2 = H2 - v + (Elev.Frm - Elev.To)) %>%
  mutate(M4 = H4 - v + (Elev.Frm - Elev.To)) %>%
  #If pole, we are looking at 100% of height instead of 80%
  mutate(MaxDepart = ifelse(Type.To == "Pole",
                            v + Height_M.To + (Elev.To - Elev.Frm), #rule for pole
                            v + 0.8*Height_M.To + (Elev.To - Elev.Frm))) %>% #rule for tree
  mutate(Depart = ifelse(Type.Frm == "Pole",
                                     ifelse((MaxDepart<Height_M.Frm), MaxDepart, Height_M.Frm), #rule for pole
                         ifelse((MaxDepart<0.8*Height_M.Frm), MaxDepart, 0.8*Height_M.Frm))) %>% #rule for tree
  mutate(Diff_maxH2 = Depart-maxH2) %>%
  mutate(Diff_maxH4 = Depart-maxH4) %>%
  mutate(Diff_M2 = ifelse(Type.To == "Pole",
                            Height_M.Frm - M2, #rule for pole
                            0.8*Height_M.Frm - M2)) %>%
  mutate(Diff_M4 = ifelse(Type.To == "Pole",
                            Height_M.Frm - M4,
                            0.8*Height_M.Frm - M4))

#### Save out
#run one at a time (n2s or s2n)
master_n2s <- master #after running n2s
master_s2n <- master #after running s2n

#bidirectional
master_bi <- rbind(master_n2s, master_s2n)

#for poles only
#master_poles <- rbind(master_n2s, master_s2n)

#before invalid flight paths were removed
# st_write(master_bi, "./Results/Flight paths/FlightPaths_2019_with Height_20220328.shp", crs=3414)

#### Remove invalid flight paths ####
#north to south
flight19_valid_n2s_H2 <- master_n2s %>%
  filter(! Diff_maxH2 < 0) %>%
  filter(! Diff_M2 < 0) %>%
  mutate(Type = ifelse(Type.To == "Pole", "Pole", 
                       ifelse(Type.Frm == "Pole", "Pole", "Tree")))

#H4 for stricter conditions
# flight19_valid_n2s_H4 <- master_n2s %>%
#   filter(! Diff_maxH4 < 0) %>%
#   filter(! Diff_M4 < 0) %>%
#   mutate(Type = ifelse(Type.To == "Pole", "Pole", 
#                        ifelse(Type.Frm == "Pole", "Pole", "Tree")))

#south to north
flight19_valid_s2n_H2 <- master_s2n %>%
  filter(! Diff_maxH2 < 0) %>%
  filter(! Diff_M2 < 0) %>%
  mutate(Type = ifelse(Type.To == "Pole", "Pole", 
                       ifelse(Type.Frm == "Pole", "Pole", "Tree")))

# flight19_valid_s2n_H4 <- master_s2n %>%
#   filter(! Diff_maxH4 < 0) %>%
#   filter(! Diff_M4 < 0) %>%
#   mutate(Type = ifelse(Type.To == "Pole", "Pole", 
#                        ifelse(Type.Frm == "Pole", "Pole", "Tree")))

#combine the two directions
flight19_valid_H2 <- rbind(flight19_valid_n2s_H2, flight19_valid_s2n_H2)
#flight19_valid_H4 <- rbind(flight19_valid_n2s_H4, flight19_valid_s2n_H4)

#poles only
# poles_valid_H2 <- rbind(flight19_valid_n2s_H2, flight19_valid_s2n_H2)

#plot to check
ggplot() +
  geom_sf(data = F1_n2s, colour = "gray") + 
  geom_sf(data = F1_s2n, colour = "gray") + 
  geom_sf(data = flight19_valid_n2s_H2, aes(colour = Type)) +
  geom_sf(data = flight19_valid_s2n_H2, aes(colour = Type))

# tm_shape(poles_valid_H2) +
#  tm_lines(col = 'blue')

### saving out valid trees ###

st_write(flight19_valid_H2, "./Results/Flight paths/FlightPaths_2019_Valid_H2.shp", crs=3414)

# st_write(poles_valid_H2, "./Results/Flight paths/FlightPaths_2019_Valid_Poles continuity_20221201.shp", crs=3414)


```

####Rasterization of vector files to heatmap

```{r Rasterize 2019 flight paths, include = F}

#variable control panel
path_MLRbuffer <- "./Working layers/P5P6_FutureMLR_50mBuffer.shp"

#load data
mlr_buffer <- st_read(path_MLRbuffer)

#remplate
raster_template = raster(extent(mlr_buffer), resolution = 0.5,
                         crs = st_crs(mlr_buffer)$proj4string)

#create buffer around flight path for allowance of Colugo to swerve
flight19_valid_n2s_buffer <- st_buffer(flight19_valid_n2s_H2, 2, endCapStyle = 'FLAT')
flight19_valid_s2n_buffer <- st_buffer(flight19_valid_s2n_H2, 2, endCapStyle = 'FLAT')
flight19_valid_H2_buffer <- st_buffer(flight19_valid_H2, 2, endCapStyle = 'FLAT')

#rasterization
flight19_n2s_raster = rasterize(flight19_valid_n2s_buffer, raster_template, field = 'Diff_maxH2', fun = 'count')
flight19_s2n_raster = rasterize(flight19_valid_s2n_buffer, raster_template, field = 'Diff_maxH2', fun = 'count')
flight19_raster_total = rasterize(flight19_valid_H2_buffer, raster_template, field = 'Diff_maxH2', fun = 'count')

#find min of values for overlapping cells
rasterlist_19 <- list(flight19_n2s_raster,flight19_s2n_raster)
stack19 <- stack(rasterlist_19)

minConnectivity_19 <- min(stack19)

#plot and see
plot(flight19_raster_total)
plot(flight19_s2n_raster)
plot(flight19_n2s_raster)

plot(minConnectivity_19)

#save out rasters
writeRaster(flight19_raster_total, "./Results/Flight paths/FlightPaths_2019_Valid_raster.tif", crs=3414)
writeRaster(flight19_s2n_raster, "./Results/Flight paths/FlightPaths_2019_Valid_raster_s2n.tif", crs=3414)
writeRaster(flight19_n2s_raster, "./Results/Flight paths/FlightPaths_2019_Valid_raster_n2s.tif", crs=3414)
# writeRaster(minConnectivity_19, "./Results/Flight paths/FlightPaths_2019_Valid_raster_effective connectivity.tif", crs=3414)

```

####Checking multi-step pathway for poles

Need to check for poles being used, does it allow for multi-step pathway i.e. south to S pole to north or north to N pole to S. If not, Colugo can travel to pole but cannot get out of it.

```{r Pole check, include = F}

path_flight19 <- "./Results/Flight paths/FlightPaths_2019_Valid_H2.shp"
path_trees2019 <- "./Working layers/TreesPoles_Remain2019_P5P6MLR_20210604.shp" #TreesPoles_2019_updated.shp
genus_exclude <- 'Elaeis|Oncosperma|Strychnos|Ziziphus|Clerodendrum|Dillenia|Calliandra'

#load tree data
trees19_sf <- st_read(path_trees2019)

#data transformation
trees19_sf2<- trees19_sf %>%
  mutate(Geom = trees19_sf$geometry) %>%
  mutate_at("MLR_Loc", as.character) %>%
  filter(! grepl(genus_exclude, Species)) %>%
  filter(! Height_M < 1)

trees19_buffer <- st_buffer(trees19_sf2, trees19_sf2$GlideRadii)

flight19_poles <- st_read((path_flight19)) %>%
  filter(Type %in% 'Pole') %>%
  mutate(Pole = "Pole")

sort(unique(flight19_poles$ID_To))
sort(unique(flight19_poles$ID_Frm))

poles_list <- c("Pole 7","Pole 8", "Pole 9")

joined_trees19_poles <- st_join(trees19_sf2, trees19_buffer, 
                          suffix = c(".To", ".Frm"), left = FALSE) %>%
                          filter(ID.To %in% poles_list | 
                                   ID.Frm %in% poles_list) %>%
                          filter(!ID.To == ID.Frm) %>%
                          filter(!Type.To == Type.Frm) %>%
                          filter(MLR_Loc.To == MLR_Loc.Frm)

flightlines19_poles <- st_sfc(mapply(function(a,b){st_cast(st_union(a,b),"LINESTRING")},
                                        joined_trees19_poles$Geom.Frm,
                                        joined_trees19_poles$Geom.To,
                                        SIMPLIFY=FALSE), crs=3414)

flight19_poles_full <- joined_trees19_poles %>%
  st_set_geometry(flightlines19_poles) %>%
  dplyr::select(-c(Lat.Frm, Long.Frm, Height_GR.Frm, Geom.Frm,
                   Lat.To, Long.To, Height_GR.To, Geom.To))

flightSamp <- flight19_poles_full %>%
  mutate(maxElev = raster::extract(x = dem19_Crop, y = flight19_poles_full, fun = max)$DEM_2019)

flightSamp <- flightSamp %>%
  mutate(diffElev_Max = maxElev-Elev.To) %>%
  mutate(LandHeight = ifelse(diffElev_Max>1.5, diffElev_Max, 1.5))

flight19 <- flightSamp

F1_n2s <- flight19 %>%
  filter(MLR_Loc.To == 'South')

F1_s2n <- flight19 %>%
  filter(MLR_Loc.To == 'North')

##run through code chunk 9 and 10 accordingly and save out "poles_valid_H2"

tm_shape(hoard) +
  tm_lines(col = 'black') +
tm_shape(flightSamp) +
  tm_lines(col = 'gray') +
tm_shape(flightSamp_same) +
  tm_lines(col = 'blue') +
tm_shape(poles_valid_H2) +
  tm_lines(col = 'red')

```

####Extracting values for Table 1 & results

```{r Extract number of paths, include = F}

#variable control panel
path_flight16 <- "./Results/Flight paths/FlightPaths_2016_Valid.shp"
path_flight19 <- "./Results/Flight paths/FlightPaths_2019_Valid_H2_20221201.shp"

#load data
flight16 <- st_read(path_flight16)
flight19 <- st_read(path_flight19)

#2016
#mean and min departure height
mean(flight16$H)
min(flight16$H)
mean(flight16$Diff_H)
median(flight16$Diff_H)
IQR(flight16$Diff_H)
min(boxplot.stats(flight16$Diff_H)$out)
max(boxplot.stats(flight16$Diff_H)$out)

#examine by species
TreeFrm16 <- flight16 %>%
  count(Spcs_Fr) %>%
  st_drop_geometry() %>%
  rename(PathFrm = n) %>%
  arrange(-PathFrm)

#count number of paths per ID
TreeFrm16 <- flight16 %>%
  count(TrID_Fr) %>%
  st_drop_geometry() %>%
  rename(PathFrm = n)

TreeTo16 <- flight16 %>%
  count(TreID_T) %>%
  st_drop_geometry() %>%
  rename(PathTo = n)

#2019
#mean and min departure height
mean(flight19$maxH2)
min(flight19$maxH2)
mean(flight19$Dff_mH2)
median(flight19$Dff_mH2)
IQR(flight19$Dff_mH2)
min(boxplot.stats(flight19$Dff_mH2)$out)
max(boxplot.stats(flight19$Dff_mH2)$out)

#examine by species
TreeFrm19 <- flight19 %>%
  count(Spcs_Fr) %>%
  st_drop_geometry() %>%
  rename(PathFrm = n) %>%
  arrange(-PathFrm)

TreeTo19 <- flight19 %>%
  count(Specs_T) %>%
  st_drop_geometry() %>%
  rename(PathTo = n) %>%
  arrange(-PathTo)

#examine by ID
TreeTo19 <- flight19 %>%
  count(ID_To) %>%
  st_drop_geometry() %>%
  rename(PathTo = n)

TreeFrm19 <- flight19 %>%
  count(ID_Frm) %>%
  st_drop_geometry() %>%
  rename(PathFrm = n) %>%
  arrange(-PathFrm)

```

####Extracting number of pixels covered with value

```{r Extract number of pixels with value, include = F}

library(exactextractr)

path_plotArea <- "./Working layers/Plot area.shp"
path_MLR2016 <- "./Working layers/0. Layers used/MLR2017.shp"
path_MLR2019 <- "./Working layers/0. Layers used/MLR2019.shp"
path_flight16_raster <- "./Results/Flight paths/FlightPaths_2016_Valid_raster.tif"
path_flight19_raster <- "./Results/Flight paths/FlightPaths_2019_Valid_raster_20221201.tif"
path_flight2022 <- "./Results/Flight paths/FlightPaths_2022_Valid.shp"
path_MLRbuffer <- "./Working layers/P5P6_FutureMLR_50mBuffer.shp"

#load data
bound <- st_read(path_plotArea)
mlr_buffer <- st_read(path_MLRbuffer)
mlr_2016 <- st_read(path_MLR2016) %>%
  st_union() %>%
  st_crop(bound)
mlr_2019 <- st_read(path_MLR2019) %>%
  st_union() %>%
  st_crop(bound)

raster16 <- raster(path_flight16_raster)
raster19 <- raster(path_flight19_raster)
flight22 <- st_read(path_flight2022) %>%
  rename(GlideFrom = MLR_L_F) %>%
  dplyr::select(c(Diff_H, geometry))
st_crs(flight22) <- 3414

load("../Colugo MS/Results/Simulation/Pole Height 12m_Norman M2/Best flight.R")
best_flight <- st_transform(best_flight, crs = 3414) %>%
  dplyr::select(Diff_H, geometry)

flight22_withpoles <- rbind(flight22, best_flight)

#generate raster for 2022 (simulation results) first
raster_template = raster(extent(mlr_buffer), resolution = 0.5,
                         crs = st_crs(mlr_buffer)$proj4string)
flight22_buffer <- st_buffer(flight22_withpoles, 2, endCapStyle = 'FLAT')
flight22_raster = rasterize(flight22_buffer, raster_template, field = 'Diff_H', fun = 'count')

flight22_buffer_withoutpoles <- st_buffer(flight22, 2, endCapStyle = 'FLAT')
flight22_raster_withoutpoles = rasterize(flight22_buffer_withoutpoles, raster_template, field = 'Diff_H', fun = 'count', progress = 'text')

#calculate the proportion of pixel within polygon with value > 1
raster16_mlr <- exact_extract(raster16, mlr_2016)
unlist(lapply(raster16_mlr, function(x) { length(which(x$value > 1)) / nrow(x) } ))

raster19_mlr <- exact_extract(raster19, mlr_2019)
unlist(lapply(raster19_mlr, function(x) { length(which(x$value > 1)) / nrow(x) } ))

raster22_mlr <- exact_extract(flight22_raster, mlr_2019)
unlist(lapply(raster22_mlr, function(x) { length(which(x$value > 1)) / nrow(x) } ))

raster22_mlr_withoutpoles <- exact_extract(flight22_raster_withoutpoles, mlr_2019)
unlist(lapply(raster22_mlr_withoutpoles, function(x) { length(which(x$value > 1)) / nrow(x) } ))

```

